---
title: "NES-LTER Chlorophyll Transect"
author: "Jaxine Wolfe"
date: "Nov 18, 2019"
output: html_document
---

## R Markdown Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# clear workspace for local development
rm(list = ls())

# assign relative path to directory
dir <- "/Users/jaxinewolfe/Documents/WHOI/NESLTER/nes-lter-chl-transect/"
# set as working directory
setwd(dir)

# define source for functions developed for the EDI packaging workflow
source("edi-utilities.R")

# install necessary libraries
# install.packages("devtools")
# install_github("EDIorg/EMLassemblyline")

# define R packages to require
libs <- c("tidyverse", "readxl", "lubridate", "devtools", "EMLassemblyline", "EML", "maps", "xml2")
# load libraries
lapply(libs, require, character.only = TRUE)
```

## Compile Chl Transect Data
```{r}

# define the list of cruises to acquire data from
cruiselist <- c("en608","en617", "en627", "en644", "ar22", "ar32")
  
# pull from API if compiled cruise data file doesnt exist in directory
if (!file.exists("nes-lter-chl-transect.csv")) {

  # compile cruise data from api
  transect_chl <- read_from_api(type = "chl", cruises = cruiselist)

  # write csv file of concatenated cruises
  write.csv(transect_chl, 'nes-lter-chl-transect.csv', row.names=FALSE)
} else {
  # read in existant data file
  transect_chl <- read_csv("nes-lter-chl-transect.csv")
}

```

## Compile SMD Chl Data

Fields to add to SMD data (this info will be supplied by the bottle summary)
• depth
• lat & lon
• datetime

Cruises that use 12-bottle rosettes will have hyphenated casts (done in rapid succession)
```{r}

# store and read in the SMD files in the directory
files <- Sys.glob("*SMD*.csv")
cruises <- lapply(files, read_csv)
# compile the data into on dataframe
smd_chl <- do.call(rbind, cruises) %>%
  filter(!is.na(cruise))

# expand SMD data
smd_expanded <- smd_chl %>%
  select(cruise, cast, niskin) %>%
  # expand cast
  mutate(cast_expand = strsplit(cast, "-")) %>%
  unnest(cast_expand) %>%
  # expand niskin
  mutate(niskin_expand = strsplit(niskin, "-")) %>%
  unnest(niskin_expand)

# assign expanded columns to numeric type
smd_expanded$cast_expand <- as.numeric(smd_expanded$cast_expand)
smd_expanded$niskin_expand <- as.numeric(smd_expanded$niskin_expand)

```

## Load CTD Bottle Summary and Metadata from API
```{r}

# compile ctd bottle summary
# pull from API if compiled cruise data file doesnt exist in directory
if (!file.exists("ctd-bottle-summary.csv")) {
  # compile the bottle summary
  ctd_summary <- read_from_api(type = "summary", cruises = cruiselist)
  # write csv file of concatenated cruises
  write.csv(ctd_summary, 'ctd-bottle-summary.csv', row.names=FALSE)
} else {
  # read in existant data file
  ctd_summary <- read_csv("ctd-bottle-summary.csv")
}

# compile ctd metadata
# pull from API if compiled cruise data file doesnt exist in directory
if (!file.exists("ctd-metadata.csv")) {
  # compile the ctd metadata
  ctd_metadata <- read_from_api(type = "metadata", cruises = cruiselist)
  # write csv file of concatenated cruises
  write.csv(ctd_metadata, 'ctd-metadata.csv', row.names=FALSE)
} else {
  # read in existant data file
  ctd_metadata <- read_csv("ctd-metadata.csv")
}


# merge depth from bottle summary using cruise-cast-bottle
# smd_metadata <- left_join(smd_expanded, ctd_summary, 
#                        by = c("cruise", 
#                               "cast_expand" = "cast", 
#                               "niskin_expand" = "niskin")) %>%
#   distinct()
# 
# 
# smd_key <- smd_metadata %>%
#   group_by(cruise, cast_expand, niskin) %>%
#   summarise(avg_depth = mean(depth))
# # this will be used in the next chunk to match transect niskin bottles
# 
# # CODE REDACTED
# # populate the key with date, lat, and lon
# smd_key$date <- Sys.time()
# smd_key$latitude <- 0
# smd_key$longitude <- 0
# 
# for (j in 1:nrow(smd_key)) {
#   # store cruise and case for given row
#   voyage <- smd_key$cruise[j]
#   cast <- smd_key$cast_expand[j]
# 
#   # subset data by cruise-cast
#   smd_subset <- smd_metadata %>% filter(cruise == voyage &
#                                         cast_expand == cast)
#   
#   # populate with info from the second row of each cruise-cast grouping
#   smd_key$date[j] <- smd_subset$date[2]
#   smd_key$latitude[j] <- smd_subset$latitude[2]
#   smd_key$longitude[j] <- smd_subset$longitude[2]
#  
# }
```

## Utility Function: Find Paired Casts

Input: df = data frame, min.time = window of minimum time difference (in mins)
 - Data needed: CTD bottle summary and metadata for casts in question
 
Output: 


```{r}

find_paired_cast <- function(df, min.time) {

# cruiselist <- unique(df$cruise)
# # read in ctd data
# ctd_metadata <- read_from_api(type = "metadata", cruises = cruiselist)
# ctd_summary <- read_from_api(type = "summary", cruises = cruiselist)

# define column to display paired cast
df$paired_cast <- NA_character_

for (j in 1:nrow(df)) {
  # store values for current row
  df_cruise <- df$cruise[j]
  df_cast <- df$cast_expand[j]

  # isolate cast metadata and bottle summary
  cast_metadata <- ctd_metadata %>% filter(cruise == df_cruise & cast == df_cast)
  cast_summary <- ctd_summary %>% filter(cruise == df_cruise & cast == df_cast)
  
  # define the min and max time for the cast 
  df_date_max <- max(cast_summary$date)
  df_date_min <- min(cast_summary$date)

  # find the cast before from the bottle summary
  cast_before <- ctd_summary %>% 
    filter(cruise == df_cruise & cast == (df_cast-1))
  # find the cast after from the ctd metadata
  cast_after <- ctd_metadata %>% 
    filter(cruise == df_cruise & cast == (df_cast+1))

  # deterine the time difference
  time_before <- min(abs(as.numeric(difftime(df_date_min, cast_before$date, units = "mins"))),
                     na.rm = TRUE)
  time_after <- min(abs(as.numeric(difftime(df_date_max, cast_after$date, units = "mins"))),
                    na.rm = TRUE)
  
  # determine paired casts if they exist
  if (time_after > min.time & time_before > min.time) {
    j <- j + 1
    next
  }
  if (time_before < min.time) {
    # isolate cast metadata to extract station
    before_metadata <- ctd_metadata %>% filter(cruise == df_cruise & cast == (df_cast-1))
    if (isFALSE(cast_metadata$nearest_station == before_metadata$nearest_station)) {
        print(paste0("paired cast found for cruise ", df_cruise, 
                     " and cast ", df_cast, ", but nearest station did not match"))
      j <- j + 1
      next
    } else {
      df$paired_cast[j] <- df_cast - 1
    }
  }
  if (time_after < min.time) {
    if (isFALSE(cast_metadata$nearest_station == cast_after$nearest_station)) {
        print(paste0("paired cast found for cruise ", df_cruise, 
                     " and cast ", df_cast, ", but nearest station did not match"))
      j <- j + 1
      next
    } else {
      df$paired_cast[j] <- df_cast + 1
    }
  }
}
return(df)
}

```

## Utility Function: Depth Window
```{r}

depth_window <- function(x) {
  y <- (2/29)*x + 2
  return(y)
}

# Example
# depth_window(50.091)
```


## Generating Bottle other method
Extract corresponding transect niskin based on depth window. Grouping by cruise and cast

```{r}

# merge geospatial and temporal metadata from bottle summary
smd_key <- left_join(smd_expanded, ctd_summary, 
                       by = c("cruise", 
                              "cast_expand" = "cast", 
                              "niskin_expand" = "niskin")) %>%
  distinct() %>%
  # find the average depth, date, lat & lon across cruise-cast-bottle
  group_by(cruise, cast_expand, cast, niskin) %>%
  summarise(depth = mean(depth),
            date = mean(date), 
            latitude = mean(latitude),
            longitude = mean(longitude))

# determine paired casts for the dataset
smd_key <- find_paired_cast(smd_key, min.time = 50)

# create row to populate with bottle other method
smd_key$bottle_other_method <- -9999

# loop through rows
for (i in 1:nrow(smd_key)) {
  # store the cruise, cast, and average depth for current row
  smd_cruise <- smd_key$cruise[i]
  smd_cast <- smd_key$cast_expand[i]
  smd_depth <- smd_key$depth[i]
  
  # subset the chl transect data by cruise
  transect_cruise <- transect_chl %>%
    select(cruise, cast, niskin, date, latitude, longitude, depth) %>%
    distinct() %>%
    filter(cruise == smd_cruise)

  # Case: the SMD cast is not found in chl transect 
  if (isFALSE(smd_cast %in% transect_cruise$cast)) {
    
    # Case: paired cast exists in chl transect
    if (smd_key$paired_cast[i] %in% transect_cruise$cast) {
      smd_cast <- smd_key$paired_cast[i]
    } else {
      print(paste0("No associated cast found for cruise ", smd_cruise, 
            ", SMD cast ", smd_cast))
      i <- i + 1
      next
    }
    
  #   # Case: no surrounding casts found in chl transect
  #   if (isFALSE((smd_cast + 1) %in% transect_cruise$cast |
  #               (smd_cast - 1) %in% transect_cruise$cast)) {
  #     print("no nearby casts")
  #     i <- i + 1
  #     next
  #   # Case: surrounding casts found in chl transect
  #   } else {
  #     # filter ctd_metadata to find the nearby sampling events
  #     actual <- ctd_metadata %>% 
  #       filter(cruise == smd_cruise & cast == smd_cast)
  #     neighbors <- ctd_metadata %>% 
  #       filter(cruise == smd_cruise) %>%
  #       filter(cast == (smd_cast+1) | cast == (smd_cast-1))
  #     
  #     # find index of the nearest station for the neighboring cast
  #     loc <- which(actual$nearest_station == neighbors$nearest_station)
  #     # reassign cast
  #     smd_cast <- neighbors$cast[loc]
  #     }
  # }
  
  # Case: the neighboring cast found does not exist in the chl transect data
  # if (isFALSE(smd_cast %in% transect_cruise$cast)) {  
  #   print("No associated cast found in chl transect data")
  #   i <- i + 1
  #   next
  # }
  
  # Case: no neighboring cast with the same nearest station was found in metadata
  # if(length(smd_cast) == 0) {
  #   print(paste0("No nearby cast found in ctd metadata for cruise ",
  #         smd_cruise, ", cast ", smd_key$cast_expand[i], ", average depth: ", smd_depth))
  #   i <- i + 1
  #   next
  }
  
  # further subset the chl transect to isolate cruise-cast data
  transect_cast <- transect_cruise %>% filter(cast == smd_cast)
  
  # find the index of the nearest depth
  ind <- which.min(abs(smd_depth - transect_cast$depth))
  
  # Print minimum depth difference if it exceeds 2m
  if (min(abs(smd_depth - transect_cast$depth)) > depth_window(smd_depth)) {
    # store min depth window
    depth <- min(abs(smd_depth - transect_cast$depth))
    # describe cruise, cast, and depth which exceeds the depth window
    print(paste0("High depth difference of ", depth, " for cruise ", 
                 smd_cruise, ", cast ", smd_key$cast_expand[i], ", average depth: ", smd_depth))
    # do not populate the bottle other method
    i <- i + 1
    next
  }
  # populate field for niskin number associated with other method
  smd_key$bottle_other_method[i] <- transect_cast$niskin[ind]
}

```


## Populating Supplementary Fields

This step will populate the fields:
  • bottle_other_method for transect_chl
  • date, lat, long, (avg) depth, and bottle_other_method for smd_chl
  • add a method of sampling column  (ethanol or acetone)
```{r}

## transect chl
transect_chl_final <- smd_key %>% ungroup() %>% 
  select(cruise, cast_expand, niskin, bottle_other_method) %>%
  left_join(transect_chl, .,
            by = c("cruise", "cast" = "cast_expand", "niskin" = "bottle_other_method")) %>%
  rename_at("niskin.y", ~"bottle_other_method")
# add method_sample col


## SMD chl
smd_chl_final <- smd_key %>% ungroup() %>% 
  select(cruise, cast, niskin, depth, date, 
         latitude, longitude, bottle_other_method) %>%
  group_by(cruise, cast, niskin, bottle_other_method) %>%
  # collapse the expanded casts
  summarise(depth = mean(depth),
            date = mean(date), 
            latitude = mean(latitude),
            longitude = mean(longitude)) %>%
  right_join(smd_chl, by = c("cruise", "cast", "niskin"))
# add method_sample col

```

## Merge Datasets
```{r}
# define headers for columns in desired order
headers <- c("cruise", "cast", "niskin", "bottle_other_method", "date", "latitude", "longitude", "depth", "replicate",
                                  "vol_filtered", "filter_size", "tau_calibration", "fd_calibration", "rb", "ra", "blank",
                                  "rb_blank", "ra_blank", "chl", "phaeo")
# reorder columns as necessary
transect_chl_final <- transect_chl_final[, headers]
smd_chl_final <- smd_chl_final[, headers]

# merge datasets by row
chl_all <- bind_rows(transect_chl_final, smd_chl_final)

# nearest station will be determined by Joe's python script?

# define numerical attributes
# num_cols <- c("tau_calibration", "fd_calibration", "rb", "ra", "blank",
#                                   "rb_blank", "ra_blank", "chl", "phaeo")
# round the nutrient values to the thousandths place
# all_chl$num_cols <- round(all_chl$num_cols, 3)

# set missing values in distance to -9999
# all_chl$mv_containing_col[is.na(all_chl$mv_containing_col)] <- -9999

# write csv file of concatenated cruises
# write.csv(chl_all, 'nes-lter-chl-transect.csv', row.names=FALSE)
```

## QA: Compare datasets

```{r}

# summarize SMD for comparison
qa_smd <- smd_key %>%
  right_join(smd_chl, by = c("cruise", "cast", "niskin")) %>%
  group_by(cruise, cast_expand, bottle_other_method, paired_cast, filter_size) %>%
  summarise(avg_chl = mean(chl, na.rm = TRUE))

# summarize transect data for comparison
qa_transect <- transect_chl_final %>%
  group_by(cruise, cast, niskin, filter_size) %>%
  summarise(avg_chl = mean(chl, na.rm = TRUE))

# define fields for logical and numerical comparisons
qa_smd$chl_compare <- FALSE
qa_smd$chl_diff <- NA_integer_

# loop through rows of SMD dataset
for (k in 1:nrow(qa_smd)) {
  # isolate variables from SMD data
  smd_cruise <- qa_smd$cruise[k]
  smd_cast <- qa_smd$cast_expand[k]
  smd_bottle <- qa_smd$bottle_other_method[k]
  smd_filter <- qa_smd$filter_size[k]
  avg_chl_smd <- round(qa_smd$avg_chl[k], 2)
  
  # subset chl transect by cruise
  qa_cruise <- qa_transect %>% filter(cruise == smd_cruise)
           
  # Case: the SMD cast is not found in chl transect 
  if (isFALSE(smd_cast %in% qa_cruise$cast)) {
    
  # Case: paired cast exists in chl transect
    if (qa_smd$paired_cast[k] %in% qa_cruise$cast) {
      smd_cast <- qa_smd$paired_cast[k]
    } else {
      print(paste0("No associated cast found for cruise ", smd_cruise, 
            ", SMD cast ", smd_cast))
      k <- k + 1
      next
    }
  }
  
  # subset chl transect by cast
  qa_cast <- qa_cruise %>%
    filter(cast == smd_cast &
             niskin == smd_bottle &
             filter_size == smd_filter)
  # isolate the mean chl from the chl transect
  avg_chl_trans <- round(qa_cast$avg_chl,2)
  
  # next row if no associated mean chl was found
  if (length(avg_chl_trans) == 0) {
    k <- k + 1
    next
  }
  
  # store difference of averages
  qa_smd$chl_diff[k] <- abs(avg_chl_smd - avg_chl_trans)
  
  # logical mean chl comparison 
  if (avg_chl_smd == avg_chl_trans) {
    qa_smd$chl_compare[k] <- TRUE
  } 
}

```


## QA: Map Sampling Locations

Call the map_locs function from edi-utility.R to map the sampling locations. Perform a visual check.

```{r}

# Map Check
map_locs(df = chl_all, xvar = "longitude", yvar = "latitude",
         region = "transect", colorvar = "cruise")

```
