---
title: "NES-LTER Chlorophyll Transect"
author: "Stace Beaulieu, Joanne Koch, Kate Morkeski, Jaxine Wolfe"
date: "`r Sys.Date()`"
output: html_document
---

## R Markdown Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# two of the required packages are installed from GitHub
# uncomment these lines to update as needed
# library(remotes)
# remotes::install_github("EDIorg/EMLassemblyline")
# remotes::install_github("WHOIGit/ediutilities")

# define R packages to require 
library(here)
library(tidyverse)
library(ediutilities)
library(readxl)
library(lubridate)
library(devtools)
library(EMLassemblyline)
library(EML)
library(maps)
library(xml2)
library(httr)

#set path to root of project
here("nes-lter-chl-transect")
```

## Compile Chl Transect Data
```{r}

base_api_url <- config::get("base_api_url")   # set in local config.yml

chl_url <- glue::glue('{base_api_url}/chl/all.csv')

api_chl <- tryCatch(read_csv(chl_url),
                 error=function(cond)
                     { message(cond)
                       return(NULL) })

# write csv file of concatenated cruises
write.csv(api_chl, 'chl-transect-api.csv', row.names=FALSE)
# if needed:
#api_chl <- read_csv('chl-transect-api.csv')

api_chl <- api_chl  %>% 
    rename(iode_quality_flag = quality_flag) %>% 
    mutate(method_contributor  = case_when(cruise == "EN608"  ~ "method_Sosik",
                                           filter_size=="<10" ~ "method_Sosik",         
                                           filter_size==">5" ~ "method_Rynearson", 
                                           filter_size==">20" ~ "method_Rynearson", 
                                           filter_size==">0" ~ "method_Sosik")) 


##### special cases for v3 ######
# AE2426 needs all TAR samples filtered out -- only a few were run so far
# filter out all of AR88 and AR95 plus AR92 TAR

# Fixed in API, saved here just in case. EN727 one TAR sample needs quality changed from NaN to 1 (cast 5) Where cast == 5, change to 1. # EN727 Sosik samples missing need flag. Where flag and rb are NA, change flag to 9 # EN727 where method is NA and filter size is 3, replace with TAR and >20; 
# api_chl <- api_chl |>
#   mutate(iode_quality_flag = case_when(str_detect(cruise, "EN727") & cast == 5 & is.na(iode_quality_flag) ~ 1, TRUE ~ iode_quality_flag)) |>
#   mutate(iode_quality_flag = case_when(str_detect(cruise, "EN727") & is.na(rb) & is.na(iode_quality_flag) ~9, TRUE ~ iode_quality_flag)) |>
#   mutate(method_contributor = case_when(str_detect(cruise, "EN727") & is.na(method_contributor) ~ "method_Rynearson", TRUE ~ method_contributor)) |>
#   mutate(filter_size = case_when(str_detect(cruise, "EN727") & str_detect(filter_size, "3") ~ ">20", TRUE ~ filter_size))

api_chl <- api_chl |>
  mutate(method_contributor = case_when(str_detect(cruise, "EN727") & is.na(method_contributor) ~ "method_Rynearson", TRUE ~ method_contributor)) 

api_chl <- api_chl |>
  mutate(filter_size = case_when(str_detect(cruise, "AE2426") & str_detect(method_contributor, "method_Rynearson") ~ "FLAG", TRUE ~ filter_size)) |>
  mutate(filter_size = case_when(str_detect(cruise, "AR92") & str_detect(method_contributor, "method_Rynearson") ~ "FLAG", TRUE ~ filter_size)) |>
  mutate(filter_size = case_when(str_detect(cruise, "AR88")  ~ "FLAG", TRUE ~ filter_size)) |>
  mutate(filter_size = case_when(str_detect(cruise, "AR95")  ~ "FLAG", TRUE ~ filter_size)) |>
  filter(filter_size != "FLAG")
  
# for testing
write.csv(api_chl, 'check-chl-transect-api.csv', row.names=FALSE)
```


## Compile Menden-Deuer Chl Data

Cruises that use 12-bottle rosettes will have hyphenated casts (done in rapid succession, considered to be paired)

```{r}

# read in the Menden-Deuer (SMD) file
smd_chl <- read_csv("NES-LTER-chla-T0-SMD-IM_flag.csv") # this version of the file has some flags edited to match growth and grazing package chl values
smd_chl <- smd_chl %>% 
  rename(tau_calibration = tau_calibrated) %>% 
  rename(fd_calibration = fd_calibrated)

smd_chl$cruise <- str_to_upper(smd_chl$cruise)

# remove apostrophes
#chlor$cast <- str_remove(chlor$cast, "'")
#chlor$niskin <- str_remove(chlor$niskin, "'")
smd_chl$niskin_second_cast <- str_remove(smd_chl$niskin_second_cast, "'")

# add niskin_second_cast to niskin
smd_chl <- smd_chl |>
  mutate(niskin_second_cast = case_when(niskin_second_cast == "NA" ~ NA, TRUE ~ niskin_second_cast)) |>
  mutate(niskin = case_when(!is.na(niskin_second_cast) ~ paste0(niskin, "|", niskin_second_cast), TRUE ~ niskin))
smd_chl <- smd_chl |> select(-niskin_second_cast)

# expand SMD data
smd_expanded <- smd_chl %>%
  select(cruise, cast, niskin) %>%
  # expand cast
  mutate(cast_expand_dash = strsplit(cast, "-" )) %>%
  unnest(cast_expand_dash) %>%
  # expand niskin
  mutate(niskin_expand_dash = strsplit(niskin, "-")) %>%
  unnest(niskin_expand_dash)

# where casts are paired, cut out second cast
smd_expanded <- smd_expanded %>%
  mutate(cast_expand_pipe = str_remove(cast, '\\|[0-9]+')) %>%
  mutate(niskin_expand_pipe = str_remove(niskin, '\\|.....')) %>%
  mutate(niskin_expand_pipe = strsplit(niskin_expand_pipe, "-")) %>%
  unnest(niskin_expand_pipe)

# merge
smd_expanded <- smd_expanded %>%
    mutate(niskin_expand = case_when(str_detect(niskin_expand_dash, "\\|") == TRUE ~ niskin_expand_pipe,
                                  TRUE ~ niskin_expand_dash)) %>%
   mutate(cast_expand = case_when(str_detect(cast_expand_dash, "\\|") == TRUE ~ cast_expand_pipe,
                                  TRUE ~ cast_expand_dash)) 
# assign expanded columns to numeric type
#smd_expanded$cast_expand <- as.numeric(smd_expanded$cast_expand)
smd_expanded$niskin_expand <- as.numeric(smd_expanded$niskin_expand)

```


## Load CTD Bottle Summary and Metadata from API
```{r}

cruises_url <- glue::glue('{base_api_url}ctd/cruises/all.csv')

all_cruises <- tryCatch(read_csv(cruises_url), 
                 error=function(cond) 
                     { message(cond)
                       return(NULL) })


# compile ctd bottle summary
# pull from API if compiled cruise data file doesnt exist in directory
if (!file.exists("ctd-bottle-summary.csv")) {
  # compile the bottle summary
  ctd_summary <- tibble()
  for (cruise_id in all_cruises$name) {
    cruise_id <- str_to_lower(cruise_id)
    url <- glue::glue('{base_api_url}ctd/bottle_summary/{cruise_id}.csv')

    btl_sum <- tryCatch(read_csv(url), 
                  error=function(cond) 
                      { message(cond)
                        return(NULL) })

   if (!is.null(btl_sum) && nrow(btl_sum) > 0) {
        btl_sum <- btl_sum %>%
            mutate(cast = as.character(cast))
        ctd_summary <- bind_rows(ctd_summary, btl_sum)
    }
  }
  # write csv file of concatenated cruises
  write.csv(ctd_summary, 'ctd-bottle-summary.csv', row.names=FALSE)
} else {
  # read in existant data file
  ctd_summary <- read_csv("ctd-bottle-summary.csv")
}


# compile ctd metadata
# pull from API if compiled cruise data file doesnt exist in directory
 if (!file.exists("ctd-metadata.csv")) {
  # compile the ctd metadata
  ctd_metadata <- tibble()
  for (cruise_id in all_cruises$name) {
    cruise_id <- str_to_lower(cruise_id)
    url <- glue::glue('{base_api_url}ctd/metadata/{cruise_id}.csv')

    metadata <- tryCatch(read_csv(url), 
                  error=function(cond) 
                      { message(cond)
                        return(NULL) })

    if (!is.null(metadata)) {
      metadata <- metadata %>%
        mutate(cast = as.character(cast))
        metadata <- metadata %>%
          mutate(
            date = as.POSIXct(date, tz = "UTC"),
            latitude = as.numeric(latitude),
            longitude = as.numeric(longitude),
            distance_km = as.numeric(distance_km)
          )

      ctd_metadata <- bind_rows(ctd_metadata, metadata)
    }
  }
   # write csv file of concatenated cruises
   write.csv(ctd_metadata, 'ctd-metadata.csv', row.names=FALSE)
 } else {
   # read in existant data file
   ctd_metadata <- read_csv("ctd-metadata.csv")
 }

# remove leading zeroes from cast
ctd_summary$cast <- str_remove(ctd_summary$cast, "^0+") 
ctd_metadata$cast <- str_remove(ctd_metadata$cast, "^0+") 

```

## Add Rynearson EN608 data

```{r}
# Rynearson data was cleaned for examples notebook and is sourced from there
tar_en608 <- read_csv("https://raw.githubusercontent.com/WHOIGit/nes-lter-examples/1df61a4627a8f845f01cd6fb05332698be3344fe/data/neslter_rynearson_chl.csv")

tar_en608 <- tar_en608 %>%
  select(-lter_station, -time_in, -time_out, -filter_type, -vol_extracted_ml, -solvent, -dilution_during_reading, -chl_cal_filename, -cal_date, -personnel_filter, -personnel_read, -fluorometer, -comments) %>%
  rename(filter_size = filter_mesh_size)

tar_en608$cast <- as.character(tar_en608$cast)

tar_en608 <- left_join(tar_en608, ctd_summary, 
                       by = c("cruise", "cast", "niskin"))

tar_en608 <- tar_en608 %>%
  select(-date.x) %>%
  rename(date = date.y)

tar_en608$method_contributor <- as.character("method_Rynearson")
tar_en608$iode_quality_flag <- as.integer(2) #quality not evaluated

# retaining df name api_chl but now includes data from API and Rynearson EN608 data not from API
api_chl$cast <- as.character(api_chl$cast)
tar_en608$cast <- as.character(tar_en608$cast)
api_chl <- bind_rows(api_chl, tar_en608)

```                           

## Determining bottle other method

This section determines depth, lat & lon, and datetime for Menden-Deuer data (merged from bottle summary).
  
```{r}

# merge geospatial and temporal metadata from bottle summary
smd_expanded <- left_join(smd_expanded, ctd_summary,
                       by = c("cruise",
                              "cast_expand" = "cast",
                              "niskin_expand" = "niskin")) 

smd_key <- smd_expanded %>%
  select(-cast_expand_dash, -cast_expand_pipe, -niskin_expand_dash, -niskin_expand_pipe) %>%
  distinct() 

smd_key <- smd_key %>%
  # find the average depth, date, lat & lon across cruise-cast-bottle
  group_by(cruise, cast, niskin, cast_expand, niskin_expand) %>%
  summarise(depth = mean(depth),
            date = mean(date),
            latitude = mean(latitude),
            longitude = mean(longitude))

```

## Join API data and Menden-Deuer data
  
  • populate the fields date, lat, long, (avg) depth for smd_chl
  • reorganize the columns
  • merge the api_chl and smd_chl datasets
  • round numerical columns
  • add project_id column 

```{r}

## SMD chl
smd_chl_final <- smd_key %>% ungroup() %>% 
  select(cruise, cast, niskin, depth, date, 
         latitude, longitude) %>%
  group_by(cruise, cast, niskin) %>%
  # collapse the expanded casts
  summarise(depth = mean(depth),
            date = mean(date), 
            latitude = mean(latitude),
            longitude = mean(longitude)) %>%
  right_join(smd_chl, by = c("cruise", "cast", "niskin"))

# add method_contributor col 
smd_chl_final$method_contributor <- "method_Menden-Deuer" 

# define headers for columns in desired order
headers <- c("cruise", "cast", "niskin", "date", "latitude", "longitude", "depth", "method_contributor", "replicate","vol_filtered", "filter_size", "tau_calibration", "fd_calibration", "rb", "ra", "blank", "rb_blank", "ra_blank", "chl", "phaeo", "iode_quality_flag")
# reorder columns as necessary
api_chl <- api_chl[, headers]
smd_chl_final <- smd_chl_final[, headers]

# merge datasets by row
api_chl$niskin <- as.character(api_chl$niskin)
chl_transect <- bind_rows(api_chl, smd_chl_final) 

```

## Format columns, add project ID

```{r}

# eliminate extra columns, set cast type, round numeric columns
chl_transect <- chl_transect %>% 
  select(-vol_filtered, -tau_calibration, -fd_calibration, -rb, -ra, -blank, -rb_blank, -ra_blank) %>% 
  mutate(cast = as.character(cast)) %>% 
  mutate(latitude = round(latitude, 4)) %>% 
  mutate(longitude = round(longitude, 4)) %>% 
  mutate(depth = round(depth, 3))  %>%
  mutate(chl = round(chl, 3)) %>% 
  mutate(phaeo = round(phaeo, 3)) 

# add project_id
chl_transect <- chl_transect  %>% 
    mutate(project_id_basic  = case_when(str_detect(cruise, "AR.*") ~ "OOI",
                                   str_detect(cruise, "EN.*") ~ "LTER",
                                   str_detect(cruise, "AE.*") ~ "LTER",
                                   str_detect(cruise, "HRS.*") ~ "LTER"
                                   ))
                           
chl_transect <- chl_transect  %>% 
    mutate(project_id_fancy  = case_when(method_contributor == "method_Menden-Deuer" ~ "LTER",
                                         method_contributor == "method_Rynearson" ~ "LTER",
                                   cruise == "AR22" ~ "JP",
                                   cruise == "AR32" ~ "JP",
                                   cruise == "AR38" ~ "JP",
                                   cruise == "AT46" ~ "LTER", 
                                   cruise == "AR77" ~ "LTER"))
    # can get more granular using method contributor if desired
chl_transect <- chl_transect  %>% 
    mutate(project_id = case_when(is.na(project_id_fancy) ~ project_id_basic,
                                  !is.na(project_id_fancy) ~ project_id_fancy))

# check for any missing method_contributor
 sum(is.na(chl_transect$method_contributor))  
                                   
```
## Join with CTD metadata to obtain nearest station and station distance

```{r}

ctd_metadata <- ctd_metadata |>
  select(cruise, cast, nearest_station, distance_km) |>
  rename(distance = distance_km) |>
  mutate(distance = round(distance, 2)) 
  
ctd_metadata$nearest_station <- replace(ctd_metadata$nearest_station, ctd_metadata$nearest_station == "MVCO-ASIT", "MVCO") 

chl_dis <- chl_transect |>
  left_join(ctd_metadata, by = c("cruise", "cast"))

```

## Read in data that are not output from API because they lack Niskin metadata 

```{r}

# samples that do not have metadata output from API

# We constructed a csv file manually with the following:
# AR24A cast 8, 4 samples: The btl file only recorded bottles 1-4. Bottle metadata for niskin 5 may be possible from CTD re-processing. Otherwise must use hand-written logsheets.
# AR24B bucket, 1 sample
# AR28B, cast 1
# AR38, 4 bucket samples
# AR39A, cast 13, niskin 7

need_metadata <-read_csv('samples_lacking_bottle_metadata-v3.csv')
need_metadata <- need_metadata %>%
  mutate(chl = round(chl, 3)) %>% 
  mutate(phaeo = round(phaeo, 3)) 

need_metadata$date <- as.POSIXct(need_metadata$date, tz = "GMT", "%m/%d/%Y %H:%M")

chl_dis$cast <- as.character(chl_dis$cast)
need_metadata$cast <- as.character(need_metadata$cast)
need_metadata$niskin <- as.character(need_metadata$niskin)

chl_dis <- chl_dis %>% anti_join(need_metadata, by = c("cruise", "cast", "niskin", "replicate", "filter_size"))

chl_qual <- bind_rows(chl_dis, need_metadata)

# check for missing quality flags
missing_qual <- chl_qual %>%
    filter(is.na(iode_quality_flag))
sum(is.na(chl_qual$iode_quality_flag)) 
sum(is.na(chl_qual$cast))
sum(is.na(chl_qual$niskin))

```

## QA: Check for samples missing nearest station

```{r}

sum(is.na(chl_qual$nearest_station))
st_dist_check <- chl_qual |> filter(is.na(nearest_station) | is.na(distance))

```
## Adjust Quality Flags, Replace Bad Data With NA, Finalize Column Names

```{r}

# replace chlorophyll and phaeo with NA for any samples flagged bad data 
chl_qual <- chl_qual %>%
  mutate(iode_quality_flag = replace(iode_quality_flag, is.na(chl), "9")) %>%
  mutate(chl = replace(chl, iode_quality_flag == "4", NA)) %>%
  mutate(phaeo = replace(phaeo, iode_quality_flag == "4", NA)) %>%
  mutate(iode_quality_flag = replace(iode_quality_flag, phaeo < 0, 3)) %>%
  mutate(iode_quality_flag = replace(iode_quality_flag, is.na(iode_quality_flag), "2"))
  
sum(is.na(chl_qual$iode_quality_flag)) 

# set chl and phaeo columns back to numeric
chl_qual$chl <- as.numeric(chl_qual$chl)
chl_qual$phaeo <- as.numeric(chl_qual$phaeo)
                              
# change some column names and properly identify the filtering regime for Sosik prefiltered samples
chl_qual <- chl_qual %>%
  rename(date_time_utc = date)  %>%
  mutate(filter_size = case_when(filter_size=="<10" ~ ">0&<10",
                                 TRUE ~ filter_size))

# eliminate milliseconds to format date correctly
chl_qual$date_time_utc <- floor_date(chl_qual$date_time_utc, unit = "seconds")

```

## Sort Columns and Write to CSV File

```{r}

#remove columns
 chl_qual <- select(chl_qual, -project_id_basic, -project_id_fancy)
 
# # add zero to casts with one digit to enable proper sorting
#  chl_qual$cast <- ifelse((nchar(chl_qual$cast)) >2, 
#                         yes = chl_qual$cast,
#                         no = str_pad(chl_qual$cast, width=2, side="left", pad="0"))
                       
#  simply sort by datetime? # puts rows without bottle metadata at end of file     
chl_qual <- chl_qual %>% arrange(date_time_utc, cruise, cast, niskin, filter_size, replicate)
names(chl_qual) 
#write.csv(chl_qual, 'nes-lter-chl-transect-rosette.csv', row.names=FALSE)
write_csv(chl_qual, 'nes-lter-chl-transect-rosette.csv', na = "NaN")

```

# Check for version differences

```{r}

 library(compareDF)
 chl_previous <- read_csv('v2/nes-lter-chl-transect-rosette.csv', show_col_types = FALSE)
 chl_previous <- chl_previous %>% arrange(date_time_utc, cruise, cast, niskin, filter_size, replicate)
 diff_check <- compare_df(chl_qual, chl_previous)
 create_output_table(diff_check, output_type='xlsx', file_name='chl_diff_check.xlsx')

```


## Compare Paired Menden-Deuer and Transect Samples

Compare the average chl for groupings of cruise, cast, bottle, and filter_size.

```{r}

# summarize SMD for comparison
qa_smd <- smd_key %>%
  right_join(smd_chl, by = c("cruise", "cast", "niskin")) %>%
  group_by(cruise, cast, niskin, filter_size) %>%
  summarise(avg_chl = mean(chl, na.rm = TRUE))

# summarize transect data for comparison
qa_transect <- api_chl %>%
  group_by(cruise, cast, niskin, filter_size) %>%
  summarise(avg_chl = mean(chl, na.rm = TRUE))
# 
# # define fields for logical and numerical comparisons
# qa_smd$chl_compare <- FALSE
# qa_smd$chl_diff <- NA_integer_

# # loop through rows of SMD dataset
# for (k in 1:nrow(qa_smd)) {
#   # isolate variables from SMD data
#   smd_cruise <- qa_smd$cruise[k]
#   smd_cast <- qa_smd$cast[k]
#   smd_bottle <- qa_smd$niskin[k]
#   smd_filter <- qa_smd$filter_size[k]
#   avg_chl_smd <- round(qa_smd$avg_chl[k], 2)
#   
#   # subset chl transect by cruise
#   qa_cruise <- qa_transect %>% filter(cruise == smd_cruise)
#            
#   # Case: the SMD cast is not found in chl transect 
#   if (isFALSE(smd_cast %in% qa_cruise$cast)) {
#     
#   # Case: paired cast exists in chl transect
#     # if (qa_smd$paired_cast[k] %in% qa_cruise$cast) {
#     #   smd_cast <- qa_smd$paired_cast[k]
#     # } else {
#     #   print(paste0("No associated cast found for cruise ", smd_cruise, 
#     #         ", SMD cast ", smd_cast))
#     #   k <- k + 1
#     #   next
#     # }
#   }
#   
#   # subset chl transect by cast
#   qa_cast <- qa_cruise %>%
#     filter(cast == smd_cast &
#              niskin == smd_bottle &
#              filter_size == smd_filter)
#   # isolate the mean chl from the chl transect
#   avg_chl_trans <- round(qa_cast$avg_chl,2)
#   
#   # next row if no associated mean chl was found
#   if (length(avg_chl_trans) == 0) {
#     k <- k + 1
#     next
#   }
#   
#   # store difference of averages
#   qa_smd$chl_diff[k] <- abs(avg_chl_smd - avg_chl_trans)
#   
#   # logical mean chl comparison 
#   if (avg_chl_smd == avg_chl_trans) {
#     qa_smd$chl_compare[k] <- TRUE
#   } 
# }

```
## Summary stats 

```{r}

# overall summary stats
qa_chl <- chl_qual %>%
  group_by(cruise, cast, niskin, filter_size) %>%
  summarise(avg_chl = mean(chl, na.rm = TRUE), 
            sd_chl = sd(chl, na.rm = TRUE), 
            min_chl = min(chl, na.rm = TRUE), 
            max_chl = max(chl, na.rm = TRUE)) 
  
# round columns 
 qa_chl$avg_chl <- round(qa_chl$avg_chl, 3) 
 qa_chl$sd_chl <- round(qa_chl$sd_chl, 3)
 
 # in order to plot averages, need to retain or join geographic data with qa_chl

```


## Plot chl values for visual inspection colored by filter_size

```{r}

# plot values vs depth
ggplot(chl_qual, aes(x=chl, y=-depth, color=filter_size))+ geom_point()

ggsave("chl-by-depth-and-filter-size.png")

# plot values vs cruise date
ggplot(chl_qual, aes(x=date_time_utc, y=chl, color=filter_size))+ geom_point()

ggsave("chl-by-date-and-filter-size.png")

# plot values vs LTER Standard Station
station <- c("MVCO", "L1", "L2", "L3", "L4", "L5", "L6", "L7", "L8", "L9", "L10", "L11", "L12", "L13", "NA")
ggplot(chl_qual, aes(x=factor(nearest_station, level = station), y=chl, color=filter_size))+ geom_point()

ggsave("chl-by-station-and-filter-size.png")

# box plots by flag
ggplot(chl_qual, aes(x=as.factor(iode_quality_flag), y=chl)) + 
   geom_boxplot() + xlab('iode_quality_flag')

ggsave("chl-by-flag.png")

ggplot(chl_qual, aes(x=as.factor(iode_quality_flag), y=chl, color=filter_size)) + 
   geom_boxplot() + xlab('iode_quality_flag')

ggsave("chl-by-flag-and-filter-size.png")

#ggplot(chl_qual, aes(x=as.factor(iode_quality_flag), y=chl, color=project_id)) + 
#   geom_boxplot() + xlab('iode_quality_flag')

ggplot(chl_qual, aes(x=as.factor(iode_quality_flag), y=phaeo)) + 
   geom_boxplot() + xlab('iode_quality_flag')

ggsave("phaeo-by-flag-and-filter-size.png")

ggplot(chl_qual, aes(x=as.factor(iode_quality_flag), y=phaeo, color=filter_size)) + 
   geom_boxplot() + xlab('iode_quality_flag')

ggsave("phaeo-by-flag.png")

```
Plot climatology

```{r}

chl_qual$doy <-as.numeric(NA)
chl_qual$year <-as.numeric(NA)

chl_qual <- chl_qual |>
  mutate(doy = yday(date_time_utc)) |>
  mutate(year = year(date_time_utc))

  # ggplot where x = cast, y = value and the lineplots are grouped by cruise
  p <- ggplot(data = chl_qual, aes(x = doy, y = chl, color = year)) +
    geom_point() +
   xlab("day of year") +
    ylab("chlorophyll concentration (µg/L)") +
    theme_minimal()
  print(p)
  ggsave("chl_doy.png")
  
   p <- ggplot(data = chl_qual, aes(x = doy, y = phaeo, color = year)) +
    geom_point() +
   xlab("day of year") +
    ylab("phaeopigment concentration (µg/L)") +
    theme_minimal()
  print(p)
  ggsave("phaeo_doy.png")

```

## QA: Map Sampling Locations

Call the map_locs function from edi-utility.R to map the sampling locations. Perform a visual check.

```{r}

# Map Check
map_locs(df = chl_qual, xvar = "longitude", yvar = "latitude",
         region = "transect", colorvar = "cruise")

# also check min/max lat/long from coverage element of final EML assembly chunk

```

## EML Assembly

This chunk outputs the final xml file for EDI through the following steps:

Step 1: Populating EML Assembly Line templates with metadata
Step 2: Calculating the geospatial and temporal coverage 
Step 3: Making the XML file 
Step 4: Inserting a custom NES-LTER parent project node 

```{r}
#chl_qual <- read.csv("nes-lter-chl-transect-rosette.csv")

# define input files
metadata <- "chl-transect-info"
edi_filename <- "nes-lter-chl-transect-rosette"
pkg_id <- "knb-lter-nes.8.3"

# Make EML Templates 
excel_to_template(here(metadata), edi_filename, rights='CC0', file_type=".md", del_rights = FALSE)

# Data Coverage
# isolate date and geospatial columns for input
date_col <- as.Date(chl_qual$date_time_utc)
lat_col <- chl_qual$latitude
lon_col <- chl_qual$longitude
# run function to determine geospatial and temporal coverage
coverage <- data_coverage(dates = date_col, lat = lat_col, lon = lon_col)

# Make EML
make_eml(path = here(),
         dataset.title = "Size-fractionated chlorophyll from water column bottle samples collected during NES-LTER Transect cruises, ongoing since 2017.",
         data.table = paste0(edi_filename, ".csv"),
         data.table.name = edi_filename,
         data.table.description = "Chlorophyll from discrete water column bottles compiled for EDI",
         temporal.coverage = c(coverage$startdate, coverage$enddate),
         geographic.description = "NES-LTER Transect",
         geographic.coordinates = c(coverage$North, coverage$East, coverage$South, coverage$West),
         maintenance.description = "ongoing",
         user.id = "NES",
         user.domain = "LTER",
         package.id = pkg_id)

# Insert Custom Project Node
project_insert(edi_pkg = pkg_id, "parent_project_NESI-II_RAPID_OOI.txt")

```


